### 2018-11-19 Vortrag/slides/sections/ddplus.tex
First version of motivation completed.

\subsection{The dd+ algorithm}
\begin{frame}{The dd+ algorithm}
	\begin{center}
		Solves it all!
	\end{center}
\end{frame}