### 2018-11-19 Vortrag/slides/sections/difficulties.tex

\subsection{Interference}
\begin{frame}{Difficulty \hashtag 1: Interference}
	\begin{itemize}
		\item $ddsimple$ works for single failure inducing-changes.
			\begin{itemize}
				\item In each recursion step it applies only the set of changes known to contain a failure inducing change.
			\end{itemize}
		\item But what if two changes exist that individually pass the tests but their combination induces failure?
	\end{itemize}

	\begin{alertblock}{Difficulty \hashtag 1: Interference}
		Let $c_1,c_2 \in \C$. $c_1$ and $c_2$ \textbf{interfere} when $test(c_1)=\cmark$, $test(c_2)=\cmark$ but $test(c_1\cup c_2) = \xmark$.
	\end{alertblock}

	\begin{exampleblock}{Idea: Leave one set of changes applied}
		If a configuration $c = c_1 \cup c_2$ with $test(c_1)=\cmark$ and $test(c_2)=\cmark$ is found by $simpledd$ then run $simpledd$ on $c_1$ while leaving $c_2$ applied and vice versa.
	\end{exampleblock}
\end{frame}

\subsection{Inconsistency}
\begin{frame}{Difficulty \hashtag 2: Inconsistency}

\end{frame}